initial_analysis_notebook

### 2021-05-11 exploratory/notebook/initial_analysis/initial_analysis_notebook.Rmd
edit lefse comparisons

  lefse_plot <- lefse_df %>% 
  	ggplot(aes(y = LDA, fill = Class, 
  			x = reorder(tax_otu_label, order))) +
  		geom_col() + 
  		theme_bw() + 
  		coord_flip() + 
  		guides(fill = 'none') + 
  		scale_fill_manual(values = c('#8de4d3','#f75b5f')) + 
  		labs(x = NULL, y = 'LDA Score (log 10)')

  abundance_plot <- shared %>% 
  	gather(OTU, abundance, matches('Otu\\d*')) %>% 
  	left_join(taxonomy, by = 'OTU') %>% 
  	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
    left_join(lefse_design, by = c('Group')) %>% 
  	group_by(Group, Genus, order, class) %>% # removed mouse_id from interactive plot
  	summarise(abundance = sum(abundance)) %>% 
  	mutate(tax_otu_label = Genus) %>% 
  	mutate(abundance = abundance / 21.07,
  		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
    #highlight_key(~mouse_id) %>% 
  	ggplot(aes(x = reorder(tax_otu_label, order), 
  			y = abundance, color = class)) + # , shape = mouse_id remove mouse_id from interactive plot
  		stat_summary(fun.data = 'median_hilow', aes(group = class),
  			fun.args = (conf.int=0.5), position = position_dodge(width = 0.5)) +
  		theme_bw() + labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
  		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
  			labels=c("0","0.1","1","10","100")) +
  		scale_color_manual(values = c('#8de4d3','#f75b5f')) + 
      scale_shape_manual(values = rep(16, 54)) + 
  		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
  		theme(panel.grid.major.y = element_blank(), 
  			panel.grid.minor = element_blank(),
  			legend.position = 'right') + 
  		coord_flip() + 
      guides(shape = 'none')

  cowplot::plot_grid(lefse_plot, 
                     abundance_plot + 
                       geom_jitter(position = position_jitterdodge(dodge.width = 0.5), alpha = 0.25) + 
                       theme(axis.ticks.y = element_blank(),
                             axis.text.y = element_blank()), 
                     nrow = 1)
}

@@

>>>>

  shared %>% 
  	gather(OTU, abundance, matches('Otu\\d*')) %>% 
  	left_join(taxonomy, by = 'OTU') %>% 
  	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
    right_join(lefse_design, by = c('Group')) %>% 
  	group_by(Group, Genus, order, class) %>% # removed mouse_id from interactive plot
  	summarise(abundance = sum(abundance)) %>% 
  	mutate(tax_otu_label = Genus) %>% 
  	mutate(abundance = abundance / 21.07,
  		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
    #highlight_key(~mouse_id) %>% 
  	ggplot(aes(x = reorder(tax_otu_label, order), 
  			y = abundance, color = class)) + # , shape = mouse_id remove mouse_id from interactive plot
  		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
      stat_summary(fun.data = 'median_hilow', aes(group = class),
  			fun.args = (conf.int=0.5), position = position_dodge(width = 0.5)) +
  		theme_bw() + 
      labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
  		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
  			labels=c("0","0.1","1","10","100")) +
  		#scale_color_manual(values = c('#8de4d3','#f75b5f')) + 
      theme(panel.grid.major.y = element_blank(), 
  			panel.grid.minor = element_blank(),
  			legend.position = 'bottom') + 
  		coord_flip() + 
      geom_jitter(position = position_jitterdodge(dodge.width = 0.5), alpha = 0.2)
}

@@

## Toxin presence for mild outcome
```{r}
plot_lefse_data('toxin_mild')
```

## Toxin presence for severe outcome
```{r fig.height = 4}
plot_lefse_data('toxin_moribund')
```

## Toxin presence and outcome severity
```{r, fig.height = 12}
plot_lefse_data('toxin_severity')
```
